---
title: "Your Mileage Might Vary"
author: "Frog"
date: "Monday, August 04, 2014"
output: html_document
---

## Executive Summary

This report analyzes the mtcars database of cars to determine if according to that database the 
transmission has a direct effect on the miles per gallon (mpg) that the car can travel. We will
examine the dataset and try to see if there is a direct relationship between mpg and the type of
transmission used in the card.

My first step would be to check for cupholders but I'd better check the correlation between mpg and the other variables in the dataset.

```{r}
library(ggplot2)
data(mtcars)
cor(mtcars$mpg,mtcars)
```

From the correlation we observe that wt (weight) is the feature with higher correlation to mpg
followed by cyl (cylinders). This makes sense as the heavier cars consume more than the light ones. I will name this observation #1 

The first thing to check is if the am feature is skewed, if there are only a few cars of either class our conclusions can be very wrong. 

```{r}
mtcars$am<-as.factor(mtcars$am)
levels(mtcars$am)<-c("auto","manual")
summary(mtcars$am)
```

With 19 automatic cars and 13 manual cars in the dataset there's no skewing.

Now I plot the mpg by the type of transmission

```{r}
qplot(mtcars$am,mtcars$mpg,geom="boxplot")
```

This plot shows that manual cars seem to have better mpg than automatic cars. I can verify this
by a simple linear model.

```{r}
reg1<-lm(mtcars$mpg~mtcars$am)
summary(reg1)
```

This model suggests that manual cars do 7.245mpg better than automatic cars.

From observation #1 now I would like to check if there is a correlation between weight and the car being automatic or not.

```{r}
cor(as.numeric(mtcars$am),mtcars$wt)
qplot(mtcars$wt,mtcars$mpg,color=mtcars$am)
```

We observe there's a strong correlation between weight and transmission. Manual cars are lighter
than automatic cars. This leads to my second model of regression trying this time both weight and
transmission as predictors for mpg.

```{r}
reg2<-lm(mtcars$mpg~as.numeric(mtcars$am) + mtcars$wt)
summary(reg2)
```

Aha! It seems that now weight has a -5.35 influence over mpg but transmission is 0.02, in other
words neutral. 

I create a third model adding the cylinders (from observation #1) to check if this holds.

```{r}
reg3<-lm(mtcars$mpg~as.numeric(mtcars$am)+mtcars$wt+mtcars$cyl)
summary(reg3)
```

It seems the weight impacts the mpg in 3.12mpg per ton, the number of cylinders also has something
to say but the transmission has a very low impact. 

Now I compare the three models using anova which curiously has nothing to do with astronomy.

```{r}
anova(reg1,reg2,reg3)
```

We observe model2 is better than 1 and 3, this doesn't mean that it is the best model for this
dataset, we could have found that using fit.step but we are trying to buy a car and not a
regression mode aren't we?

Since the second model seems to be the better one, let's diagnose its residuals.

Model 2: mpg ~ wt
```{r}
par(mfrow=c(2,2))
plot(reg2)
```

Outliers! the world would be much better without them. We can see that even considering outliers there's a good fit for the residuals meaning our model is accurate.

## Conclusion

With this dataset we can't conclude if they type of transmission affects the fuel consumption of
a car, lighter cars have better mpg than heavier cars. Manual transmission cars tend to be lighter
than automatic cars and that's why they show better fuel consumption. 

Buy a light car it doesn't matter if it has a stick or not.



